Responsible Machine Learning Techniques\\\large{Interpretable Models, Post-hoc Explanation, and Discrimination Testing}

### 2019-12-20 article-information-2019.tex
add last section

\subsection{Impact of Discrimination Testing on Model Use and Adoption}

\subsection{Viable Discrimination Remediation Approaches}\label{ssec:disc_rem}

@@

The black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the security vulnerabilities inherent in ML are each serious and difficult problems on their own. However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g., the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}. High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}. This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{\textcolor{red}{24}}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems. Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena. Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML. However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks, and that explanations can mask ML discrimination. Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory. Accuracy, discrimination, and security characteristics of a system can change over time as well. Simply testing for these problems at training time, as presented in Sections \ref{sec:m_and_m} and \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems. Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.

>>>>

\subsection{Discrimination Testing and Remediation in Practice}\label{ssec:disc_rem}

The heightened interest in exploring and fixing algorithmic bias has led to a significant body of work focused in using ML to diminish discrimination in model outcomes \cite{friedler2019comparative}. Broadly, these can be defined into two groups: more traditional methods that mitigate discrimination by searching across possible algorithmic and feature specifications, and more recently developed approaches that change the algorithms or input data themselves in order to mitigate disparities. Many approaches that have been developed in the last five to seven years focus on altering the algorithm itself, preprocessing the data, or post-processing the predictions in order to diminish class-control correlations or dependencies. IBM’s \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} toolkit provides an open source Python package for many of the metrics and techniques that have been developed recently.

Whether these methods are suitable for a particular use case depends on the legal environment in which the model will be used and the use case itself.  Great care must be taken in order to ensure that the appropriate fairness metrics are chosen, because certain metrics may not be appropriate for some use cases. Additionally, the effects of changing the model must be viewed holistically.  For example, the mortgage data disparity analysis in Section \ref{ssec:dis_mort} shows that if one were to choose $g^\text{MGBM}$ over $g^\text{XNN}$ because $g^\text{MGBM}$ has a lower FPR ratio for blacks, it would ultimately lead to a higher FPR for blacks overall, which may represent doing more harm than good.

Furthermore, using some recently developed discrimination mitigation methods may lead to non-compliance with anti-discrimination laws and regulations.  A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.''\footnote{As discussed in \textit{Bazemore v. Friday}, 478 US 385 (1986), a discrimination in pay case, the US Supreme Court wrote that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.} A model developed without inclusion of class status (or proxies thereof) usually considers similarly situated people the same on the dimensions included in the model: people who have the same feature values will usually have the same model output.  Obviously, the inclusion of protected class status will change model output by class.  With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal.  Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.

Of the newer class of fairness enhancing interventions in ML, within-algorithm discrimination mitigation techniques that do not use class information may be more likely to be acceptable in highly regulated settings today. These techniques work by incorporating a loss function where more discriminatory paths or weights are penalized and will only be used by the model if the increase in model performance overcomes the penalty (the relative level of performance-to-discrimination penalty is determined via a hyperparameter choice). Other mitigation strategies that only alter hyperparameters or algorithm choice are also likely to be acceptable. And feature selection techniques that have been used in traditional modeling (ordinary least squares, logistic, and simple decision tree models) are likely to continue to be accepted in regulatory environments.  For further discussion of potential and utilized techniques that can mitigate disparate impact in financial services, see \citet{schmidt2019introduction}.

Regardless of the methodology chosen to minimize disparities, advances in computing have enhanced the ability to search for less discriminatory models.  Prior to these advances, only a small number of alternative algorithms could be tested for lower levels of disparity without causing infeasible delays in model implementation.  Now, tens of thousands of alternative models or more can be quickly tested for lower discrimination and better predictive quality.  An additional opportunity arises as a result of ML itself: the well-known Rashomon effect, or the multiplicity of good ML models for most datasets. Thus, it is now feasible to train more models, find more good models, and test more models for discrimination, and among all those tested models, there are likely to be some with high predictive performance and low discrimination.

@@

The black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the security vulnerabilities inherent in ML are each serious and difficult problems on their own. However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g., the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}. High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}. This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{25}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems. Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena. Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML. However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks, and that explanations can mask ML discrimination. Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory. Accuracy, discrimination, and security characteristics of a system can change over time as well. Simply testing for these problems at training time, as presented in Sections \ref{sec:m_and_m} and \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems. Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.
